---
title: "Data 607 Project 1"
author: "Jiaxin Zheng"
date: "2024-10-09"
output:
  pdf_document: default
  html_document: default
---
Introduction:

In this project we take chess tournament results text file into a .csv file. Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605


Data Cleaning Strategy:
1. read the data from text file
2. removing all the |’s , -’s and NA columns
3. separate the rows into more clean and combine as a readable data table
4. separate the  USCF ID / Rtg (Pre->Post)
5. Calculating the opposing player rating


```{r loading the package}
library(tidyverse)
```

```{r load the tournament data}

url <- "https://raw.githubusercontent.com/Jennyjjxxzz/Data-607_Project1/refs/heads/main/tournamentinfo.txt.txt"
tournament <- read.delim(url, header = FALSE, sep = "\n")

```

```{r read the data}
str(tournament)
head(tournament)
```
```{r clean and filter the data}

cols <- c('player_num','name','total_pts','round_1','round_2','round_3','round_4',
          'round_5','round_6','round_7','NA')
```

```{r remove the dashes and "|"}

df_tournament <- read.csv(url, sep="|", header = FALSE, skip = 3, col.names = cols)
dashes <- "-----------------------------------------------------------------------------------------"
df_tournament <- df_tournament %>% filter(player_num != dashes)

```


```{r separate the rows into more clean and combine as a readable data table}
df1 <- df_tournament %>%
  filter(row_number() %% 2 == 1)

```

```{r}
df2 <- df_tournament%>%
  filter(row_number() %% 2 != 1)
```

```{r combine the two dataframes and remove the NA columns}

combine_df<- cbind(df1, df2)
head(combine_df)

```
```{r clean the data, and rename}

combine_df<- combine_df %>%
  subset(select=c(1:10, 12:13))

colnames(combine_df)<- c("Player_num", "Name", "Total_Points", "Round1", "Round2", "Round3", "Round4", "Round5", "Round6", "Round7", "State", "Opponent_Info")
view(combine_df)

```


```{r Extract the Pre-Rating and extract the Post-Rating}
combine_df <- combine_df %>%
  mutate(Pre_Rating = str_extract(Opponent_Info, "(?<=R: )\\d+"),  # Extract the Pre-Rating
         Post_Rating = str_extract(Opponent_Info, "(?<=->)\\d+"))  # Extract the Post-Rating


combine_df <- combine_df %>%
  rowwise() %>%
  mutate(Average_Opponent_Rating = mean(as.numeric(c(str_extract(Round1, "(?<=R: )\\d+"),
                                                     str_extract(Round2, "(?<=R: )\\d+"),
                                                     str_extract(Round3, "(?<=R: )\\d+"),
                                                     str_extract(Round4, "(?<=R: )\\d+"),
                                                     str_extract(Round5, "(?<=R: )\\d+"),
                                                     str_extract(Round6, "(?<=R: )\\d+"),
                                                     str_extract(Round7, "(?<=R: )\\d+"))), na.rm = TRUE))

#columns for the cleaned dataframe
final_df <- combine_df %>%
  select(Name, State, Total_Points, Pre_Rating, Average_Opponent_Rating)

view(final_df)
head(final_df)

```

```{r}
write.csv(final_df, "tournament_cleaned.csv", row.names = FALSE)
```






